---
title: "class09"
author: "Sharon Tu"
date: "February 7, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

## Read our data
```{r}
read.csv("WisconsinCancer.csv")
```

```{r}
wisc.df <- read.csv("WisconsinCancer.csv")
```

```{r}
# Save your input data file into a new "data" directory
fna.data <- "WisconsinCancer.csv"

#Complete the following code to input the data and store as wisc.df
wisc.df <- read.csv(fna.data)
```

# Let's see the first 6 rows of data
```{r}
head(wisc.df)
```

How many samples (i.e. patients) are in this data-set?
```{r}
nrow(wisc.df)
```

```{r}
colnames(wisc.df)
```

Next use as.matrix() to convert the other features (i.e. columns) of the data (in columns 3 through 32) to a matrix. Store this in a variable called wisc.data.

```{r}
# Convert the features of the data: wisc.data
wisc.data <- as.matrix(wisc.df[,3:32])
```

Add rownames to our new matrix of data
```{r}
rownames(wisc.data) <- wisc.df$id
```

# Create diagnosis vector by completing the missing code
```{r}
diagnosis <- as.numeric(wisc.df$diagnosis == "M")
```

How many cancer (M) and non cancer samples do we have in our dataset?
```{r}
table(wisc.df$diagnosis)
```

```{r}
#diagnosis <- as.numeric(wisc.df$diagnosis)
#tmp <- rep(0, nrow(wisc.df))
wisc.df$diagnosis =="M"
as.numeric(wisc.df$diagnosis == "M")
```

#Let's double check
```{r}
cbind(diagnosis, wisc.df$diagnosis)
```

Q1. How many observations are in this dataset?
```{r}
nrow(wisc.data)
```

Q2. How many variables/features in the data are suffixed with _mean?
```{r}
colnames(wisc.data)
```
```{r}
inds <- grep("_mean", colnames(wisc.data))
length(inds)
```


Q3. How many of the observations have a malignant diagnosis?
```{r}
sum(diagnosis)
```

# Let's do PCA
```{r}
#Perform PCA on wisc.data by completing the following code
wisc.pr <- prcomp(wisc.data, scale=TRUE)
summary(wisc.pr)
```

Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)?



Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?
Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?

## Let's see the plots of our results
First: A standard biplot (not rly helpful here)
```{r}
biplot(wisc.pr)
```

We need to make our own PCA plot
```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2])
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=diagnosis+1, xlab = "PC1", ylab = "PC2")
```

## Variance captured in each PC

This info is in the $sdev component of our PCA result
```{r}
variance <- wisc.pr$sdev^2
sum(variance)

#Variance explained by each principal component: PVE
round((variance/sum(variance)) * 100, 1)
pve <- round((variance/sum(variance)) * 100, 1)
```

#Plot pve
```{r}
plot(pve, type = "o")
```

```{r}
barplot(pve)
```

# make the plot a bit fancier
```{r}
# Alternative scree plot of the same data, note data driven y-axis
barplot(pve, ylab = "Precent of Variance Explained",
     names.arg=paste0("PC",1:length(pve)), las=2, axes = FALSE)
axis(2, at=pve, labels=round(pve,2))

#
```

##Hierarchical clustering of cancer case data

For hierarchical clustering, we need a few things:
1. the distance matrix **dist()** function
2. the **hclust()** function
3. use the **cutree()** function to find cluster membership vector

#Scale the wisc.data data: data.scaled
```{r}
data.scaled <- scale(wisc.data)
```

```{r}
apply(wisc.data, 2, mean)
round(apply(wisc.data, 2, sd), 1)

```

```{r}
round(apply(data.scaled, 2, sd), 1)
```

Looks like we need scaling!

Calculate the (Euclidean) distances between all pairs of observations in the new scaled dataset and assign the result to  data.dist.

```{r}
data.dist <- dist(data.scaled)
```

```{r}
wisc.hclust <- hclust(data.dist, method = "complete")
```

```{r}
plot(wisc.hclust)
abline(h=19, col="red", lty=2)
```


## Cluster in PCA space

For clustering, we need:
1. Distance matrix
2. clustering function
3. Cutree

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=diagnosis+1)
```

```{r}
pc.dist <- dist(wisc.pr$x[,1:2])
pc.hclust <- hclust(pc.dist, method="ward.D2")
plot(pc.hclust)
```

```{r}
grps3 <- cutree(pc.hclust, k=3)
table(grps3)
```

```{r}
table(grps3, diagnosis)
```

## Predictions
We will use the predict() function that will take our PCA model from before and new cancer cell data and project that data onto our PCA space.

```{r}
#url <- "new_samples.csv"
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```


```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=grps3)
points(npc[,1], npc[,2], col="blue", pch=16)
```

```{r}
plot(wisc.pr$x[,1:2], col=grps3)
```

```{r}
plot(wisc.pr$x[,1:2], col=diagnosis+1)
points(npc[,1], npc[,2], col="blue", pch=15, cex=3)
```



