---
title: "Class 14: Genome Informatics 1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Asthma SNPs in MXL population

Read file from Ensembl

```{r}
mxl <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378 (1).csv")
head(mxl)
```

How many GIG genotypes are in this set?

```{r}
table(mxl$Genotype..forward.strand.)
```

```{r}
sum(mxl$Genotype..forward.strand.=="G|G")
```

Percent values

```{r}
table(mxl$Genotype..forward.strand.) / nrow(mxl) * 100
```

## RNA-Seq

Fastq quality scores

Q. Does the first sequence have good quality?

```{r}
#install.packages("seqinr")
#install.packages("gtools")
library(seqinr)
library(gtools)
q <- "DDDDCDEDCDDDDBBDDDCC@"
asc(s2c(q)) - 33
```


## Population scale analysis

Does the genotype for rs9067378 influence the expression of the important asthma linked genes we found?

```{r}
geno <- read.table("rs8067378_ENSG00000172057.6.txt")
head(geno)
```

```{r}
table(geno$geno)
```

```{r}
inds <- geno$geno == "G/G"
summary(geno$exp[inds])
```

```{r}
inds.aa <- geno$geno == "A/A"
summary(geno$exp[inds.aa])
```

```{r}
inds.ag <- geno$geno == "A/G"
summary(geno$exp[inds.ag])
```

Some nice figure that shows the distribution of these expression value...?

```{r}
boxplot(exp ~ geno, geno)
```

```{r}
# Generate a boxplot for expression data of different genotype groups
p <- boxplot(exp~geno, data=expr, xlab="rs8067378 genotype", ylab="ENSG00000172057.4 (RPKM)", notch=T)
```


```{r}
## Histogram of the exp column with ggplot2
ggplot(expr, aes(exp, fill = geno)) + geom_density(alpha = 0.2)
```

```{r}
# Boxplot with the data shown
ggplot(expr, aes(geno, exp, fill=geno)) + 
  geom_boxplot(notch=TRUE, outlier.shape = NA) + 
  geom_jitter(shape=16, position=position_jitter(0.2), alpha=0.4)
```

